---
title: "Life_Table"
author: "Zhuodiao Kuang"
date: "`r Sys.Date()`"
output: github_document
---



# Load packages

```{r , load_packages, message = FALSE,warning=FALSE}
library(survival)
library(tidyverse)
library(ggfortify)
library(dplyr)
library(ggplot2)
library(biostat3)
library(knitr)
```

```{r setup, echo = FALSE, cache = FALSE, message = FALSE}
rm(list = ls())

## Global options
opts_chunk$set(echo      = TRUE,
	             cache     = TRUE,
               prompt    = FALSE,
               tidy      = FALSE,
               comment   = NA,
               message   = FALSE,
               warning   = FALSE,
               dpi       = 150,
               fig.width = 4, 
               fig.asp   = 5/7,
               fig.align = "center")
# attr.source = '.numberLines'
```

# Ovarian Cancer:

* futime: survival or censoring time(day)
* fustat: censoring status(censor = 0)
* age: in years
* resid.ds: residual disease present(1=no, 2=yes)
* rx: treatment group
* ecog.ps: ECOG performance status(1 is better)

```{r ovarian, message=FALSE, warning=FALSE}
data("ovarian")
attach(ovarian)
```


# Life-table summary stratified by rx
```{r}
res <- summary( survfit( Surv(futime, fustat)~rx, data=ovarian))
cols <- lapply(c(2:6, 8:11) , function(x) res[x])
tbl <- do.call(data.frame, cols)
tbl
```


# Create life-table stratified by rx

```{r}
ovarian_rx1 <- ovarian |>
  filter(rx == 1) |>
  arrange(futime)

ovarian_rx2<- ovarian |>
  filter(rx == 2)|>
  arrange(futime)

lifet1<-lifetab2(Surv(futime, fustat == 1)~1,ovarian_rx1)
lifet2<-lifetab2(Surv(futime, fustat == 1)~1,ovarian_rx2)
print(lifet1, digits = 2)
print(lifet2, digits = 2)
```


# Plot hazard function by rx based on life-table estimate

```{r}
hazard1<-lifet1 |>
  dplyr::select(tstart, tstop, hazard) |>
  mutate(tmedian = (tstart+tstop)/2, rx ="1")

hazard2<-lifet2 |>
  dplyr::select(tstart, tstop, hazard) |>
  mutate(tmedian = (tstart+tstop)/2, rx ="2")

hazard <- rbind(hazard1,hazard2)
```

```{r}
ggplot(hazard, aes(x = tmedian, y = hazard, color = rx)) +
  geom_point()+
  geom_line()


```



# Plot K-M survival function by rx

```{r}
ovarian.survfit <-
  survfit(Surv(futime, fustat)~rx,data= ovarian)

ovarian.survfit |>
  autoplot() +
  ylab("S(t)") +
  xlab("Time")
```

# Median survival time for each treatment group

For the group 1(rx = 1), the median survival time is $534.5(\frac{431+638}{2})$ days.
For the group 2(rx = 2), the median survival time is not sure, because over half of patients are still censored.

# Compare survival function estimations between K-M and F-H methods



